Hierarchical wrinkling in a confined permeable biogel

### 2015-04-27 fig_plis/article_SI.tex
Déplace Wavelength measurements au début des SI
Réorgnise les SI Typos et bémols

To form a gel we start from an aqueous dispersion of caseins and GDL. GDL slowly acidify the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}. While still liquid, this yoghurt-forming suspension is injected into an home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a. The cell is sealed an no external stimulus is applied during gelation or pattern formation. Both top and bottom walls are coated by acrylamide brushes, effectively turning off the casein adhesion so that the gel is only anchored to the remaining four sides of the cell.

@@

We do not observe coarsening of the patterns as in~\cite{Yoo2003}. We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}. Indeed when the amplitude saturates the gel flattens against the walls and create sharp edges. To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:ripples}. The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening.

@@

Based those observations, we examine possible scenarios upon shrinking and swelling of the gel. As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c. For the swelling back we consider three options:

@@

\noindent
(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film. This simple buckling situation leads to a primary pattern composed of a single bump of size $L$. As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}. That is not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.

\noindent
(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom wall. To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths. The transverse load could be gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirm that it plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e. with the weight acting longitudinally rather than transversely. Furthermore, the transverse load cannot be adhesion~\cite{Vella2009a} since the film is not adhering to the top and bottom walls. Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample. Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}.

@@

@@

For each experiment we measure independently $\lambda$ and all the parameter that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}). $\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Methods and Supplementary methods.

@@

Despite changing the gel properties and the cell geometry we always observe wrinkling proving the robustness of this approach to form wrinkling patterns. We are in the position to vary $H/H^\star$ from 0.1 to 10 so that we can test the crossover between Darcy and Poiseuille mechanism. In Figure~\ref{fig:DarcyPoiseuille}a and b we have plotted the measure wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively. If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$. Indeed, both model seems to apply in their domain of validity albeit with a prefactor. The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data. To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods. The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.

@@

In conclusion we have taken advantage of the properties of the casein protein to engineer and study the wrinkling of a porous confined biogel. In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of continuous acidification. Together with carefully controlled adhesion of the gels to the boundaries, this route to form permeable biogels is the key to produce an original cascade of wrinkling patterns. We developed a model that captures the dynamical origin of the constraints exerted on the gel and predicts the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall. Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins. As such it illustrates daily life problems such as wrinkles encountered while putting up wall paper and could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}.

>>>>

To form a gel we start from an aqueous dispersion of caseins and glucono-$\delta$-lactone (GDL). GDL slowly acidify the medium and triggers gelation of the caseins~\cite{Roefs1986,Bremer1989,Lucey1998,Dickinson2002}. While still liquid, this yoghurt-forming suspension is injected into an home-made optical cell, as sketched in Figure~\ref{fig:dynamics}a. The cell is sealed and no external stimulus is applied during gelation or pattern formation. Both top and bottom walls are coated by acrylamide brushes, effectively turning off the casein adhesion so that the gel is only anchored to the remaining four sides of the cell.

@@

XXX formulation XXX We do not observe coarsening of the patterns as in~\cite{Yoo2003}. We believe the wrinkling pattern is stable against coarsening because of the bending rigidity of the gel~\cite{LeGoff2014}. Indeed when the amplitude saturates the gel flattens against the walls and create sharp edges. To accommodate those sharp edges with finite bending rigidity, the flat part of the gel actually ripples, see Supplementary Figure~\ref{fig:ripples}. The ripples created by opposite sides of a same flat patch interfere, creating barriers against coarsening.

@@

Based on those observations, we examine possible scenarios upon shrinking and swelling of the gel. As the gel forms it expels the solvent and detaches from top and bottom wall, see Figure~\ref{fig:dilation}b and c. For the swelling back we consider three options:

@@

\noindent
(ii) In Figure~\ref{fig:dilation}e, we assume that nothing hinders the buckling of the gel film. This simple buckling situation leads to a primary pattern composed of a single bump of size $L$. As the bump amplitude grows larger saturates on the top and bottom wall, buckling of smaller wavelength may appear, as observed by~\cite{Roman1999}. That is also not what we observe since our primary wavelength $\lambda$ is much smaller that $L$.

\noindent
(iii) We are left with the case of Figure~\ref{fig:dilation}f where a wrinkling pattern emerges spontaneously with a wavelength $\lambda\ll L$ selected before saturation on the top and bottom walls. To obtain such an instability, one must balance the bending rigidity of the gel film with a transverse load $\sigma_{\perp}$ that hinders buckling on large wavelengths. The transverse load could be gravity~\cite{Smoluchowski1910, Kolinski2009, Vella2009, Pineirua2013, Lucantonio2013}, however we confirm that it plays no significant role by successfully repeating the wrinkling pattern in a cell held vertically, i.e. with the weight acting longitudinally rather than transversely. Furthermore, the transverse load cannot be adhesion~\cite{Vella2009a} since the film is not adhering to the top and bottom walls. Influence of the side boundaries~\cite{Vandeparre2011a, Li2013} is also excluded since the wavelength of the pattern is uniform throughout the sample. Finally as the four side boundaries remain fixed $\sigma_{\perp}$ cannot come from the resistance to uniaxial stretching~\cite{Cerda2003}. XXX conclure plus clairement XXX

@@

XXX repetition entre paragraphe d'avant et d'après XXX

@@

For each experiment we measure independently $\lambda$ and all the parameters that go into the expressions of $\lambda_D$, Equation~(\ref{eq:prefDarcy}) and $\lambda_P$, Equation~(\ref{eq:prefPoiseuille}). $\lambda$ is measured by optical microscopy; $e$, $h$, $H_1$, $H_2$, $\tau$ are measured by confocal microscopy; $\alpha$ is measured via capillary experiment and $B$ by rheology, see Methods and Supplementary methods.

@@

Despite changing the gel properties and the cell geometry we always observe wrinkling proving the robustness of this approach to form wrinkling patterns. We are in the position to vary $H/H^\star$ from 0.1 to 10 so that we can test the crossover between Darcy and Poiseuille mechanism. In Figure~\ref{fig:DarcyPoiseuille}a and b we have plotted the measure wavelength as a function of the model prediction $\lambda_D$ and $\lambda_P$ respectively. If the predictions are correct we expect $\lambda_\text{exp}=\lambda_D$ when $H<H^*$ and $\lambda_\text{exp}=\lambda_P$ when $H>H^*$. Indeed, both model seems to apply in their domain of validity albeit with a prefactor XXX Thomas dit pas clair XXX. The failure of the Poiseuille model is only visible at small wavelengths where an unexplained offset (dashed line in Figure~\ref{fig:DarcyPoiseuille}b) is necessary to fit the data. To go further, we derive a complete model summing up the influence of both Poiseuille and Darcy dissipation, see Supplementary Methods. The results of this mixed model are displayed in Figure~\ref{fig:DarcyPoiseuille}c and agree with $\lambda_\text{exp}$ for the whole $H/H^*$ range.

@@

In conclusion we have taken advantage of the properties of the casein protein to engineer and study the wrinkling of a porous confined biogel. In particular, we have fine-tuned the interplay between the attraction and the repulsion at the protein level to assemble dense glassy networks that shrink and swell as a result of continuous acidification. Together with carefully controlled adhesion of the gels to the boundaries, this route to form permeable biogels is the key to produce an original cascade of wrinkling patterns. We developed a model that captures the dynamical origin of the constraints exerted on the gel and predicts the wrinkling wavelength: Poiseuille flows of the solvent above and below the gel when the film is far from any wall, Darcy flow of the solvent through the gel when the film is close to a wall. Such a model experiment, because it relies on pH-induced charge stabilisation/destabilisation, should be generalisable to different kinds of proteins. As such it could set a benchmark to explore potential applications in micro-fingerprinting~\cite{Bae2015} or in optical devices such as diffraction gratings or Fresnel lenses~\cite{Li2013}, and the possibility of wrinkling in confined porous soft materials immersed in a buoyancy-matched viscous medium such as biological tissues~\cite{Ma2004,Longley2013,Fleury2015}. XXX texture food XXX
